---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Philipp von Praun (12405779)"
date: "`r Sys.Date()`"
output: distill::distill_article
---

## General Remarks

-   You can submit your solutions in teams of up to 3 students.
-   Include all your team-member's names and student numbers (Martrikelnummern) in the `authors` field.
-   Please use the exercise template document to work on and submit your results.
-   Use a level 2 heading for each new exercise and answer each subtask next to it's bullet point or use a new level 3 heading if you want.
-   Always render the R code for your solutions and make sure to include the resulting data in your rendered document.
    -   Make sure to not print more than 10 rows of data (unless specifically instructed to).
-   Always submit both the rendered document(s) as well as your source Rmarkdown document. Submit the files separately on moodle, **not** as a zip archive.

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it in git.

a)  Start by initializing a new R project with git support, called `2024-exeRcise-sheet-3`. If you forgot how to do this, you can follow this [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).
b)  Commit the files generated by Rstudio.
c)  For all of the following tasks in this exercise sheet we ask you to always commit your changes after finishing each subtask e.g. create a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their answer. If you complete tasks in a different order or forget to commit one, this is no problem. If you change your answers you can just create multiple commits to track the changes.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a commit after this answer, see *1c*)

    Pros: - Git is great for collaborating
          - Git allows you to go back in time and understand the stages a certain project went through
          
    Cons: - Git has to be explicitly used, which means it does not run in the                   background/automatically
          - Git can be quite complicated to operate, especially at the beginning

e)  Knit this exercise sheet. Some new files will automatically be generated when knitting the sheet e.g. the HTML page. Ignore these files, as we only want to track the source files themselves.

## Exercise 2: Putting your Repository on GitHub (3.5 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named `exeRcise-sheet-3`. Make sure you create a **public repository** so we are able to see it for grading. Add the link to the repository below:

  https://github.com/Loirinho1905/exeRcise-sheet-3.git

b)  Push your code to this new repository by copying and executing the snippet on github listed under `…or push an existing repository from the command line`.

c)  Regularly push your latest changes to GitHub again and especially do so when you are finished with this sheet.

## Exercise 3: Baby-Names in Munich (3.5 Points)

Download the latest open datasets on given names ("Vornamen") from the open data repository of the city of Munich for the years `2023` and `2022`.

Link: <https://opendata.muenchen.de/dataset/vornamen-von-neugeborenen>

a)  Download the data for both years and track it in git. For small datasets like these adding them to git is not a problem.

b)  Load the data for both years into R. Check the type of the count variable ("Anzahl") and look into the data to determine why it is not numeric? Fix the problem in an appropriate manner, it is OK if some of the counts are inaccurate because of this. Explain your solution and the repercussions.

```{r}
vornahmen2023 <- read.csv("data/Vornamen2023.csv")
vornahmen2022 <- read.csv("data/Vornamen2022.csv")
```

```{r}
is.numeric(vornahmen2023$Anzahl)
typeof(vornahmen2023$Anzahl)
is.numeric(vornahmen2022$Anzahl)
typeof(vornahmen2022$Anzahl)
```

The variable "Anzahl is a character variable, because it does not include accurate counts beneath four, listing these cases as "4 or less" instead.

```{r}
library(tidyverse)

vornahmen22num <- vornahmen2022 %>% 
  mutate(Anzahl = case_when(Anzahl != "4 oder weniger" ~ Anzahl, 
                            Anzahl == "4 oder weniger" ~ "2.5"))
vornahmen22num$Anzahl <- as.numeric(vornahmen22num$Anzahl)

vornahmen23num <- vornahmen2023 %>% 
  mutate(Anzahl = case_when(Anzahl != "4 oder weniger" ~ Anzahl, 
                            Anzahl == "4 oder weniger" ~ "2.5"))
vornahmen23num$Anzahl <- as.numeric(vornahmen23num$Anzahl)
```

My solution is a pragmatic one. Since the group of "4 oder weniger" includes a huge amount of names, it would not do to exclude them from the dataset. As we cannot know the exact count of these names, however, we have to account for the variance in their numbers. This is why my solution opts to assign them the value 2.5, which, assuming a balanced distribution, is the median between 1 and 4. It is therefore a compromise between the imprecision and completeness of the data in question.

c)  Calculate the total number of babies born in Munich in 2022 and 2023. Which year had the bigger baby-boom?

```{r}
sum(vornahmen22num$Anzahl)
sum(vornahmen23num$Anzahl)
```

More babies were born in Munich in 2022 than in 2023.

d)  Add a new column `year` to both datasets which holds the correct year for each.

```{r}
vornahmen22_year <- vornahmen22num %>% mutate(year = 2022)
vornahmen23_year <- vornahmen23num %>% mutate(year = 2023)
```

e)  Combine both datasets into one using `bind_rows()`.

```{r}
vornahmen2223 <- bind_rows(vornahmen22_year, vornahmen23_year)
```

f)  Combine the counts for same names to determine the most popular names across both years. Print out the top 10 names in a nicely formatted table for both years. Include a table caption.

```{r}
library(knitr)

vornahmen_comb <- vornahmen2223 %>% group_by(Vorname) %>% 
  summarise(Anzahl = sum(Anzahl)) %>% arrange(desc(Anzahl))

vornahmen_comb %>% head(10) %>% kable(
  caption = "Top 10 most popular baby names in Munich (Germany) in 2022 and 2023, combined"
  )
```


## Exercise 4: Open Analysis (4 points)

This exercise is a bit more open-ended. You can choose any dataset from [Our World in Data](https://ourworldindata.org/) and analyze it, while determining the research question yourself.

a)  Go to <https://github.com/owid/owid-datasets/tree/master/datasets> and choose a dataset that interests you. You can have a look at <https://ourworldindata.org/> to gather some inspiration.

b)  Download the dataset and track it in git.

c)  Put the name / title of the dataset and a link to it below.

-   Dataset Name: Who Americans spend time with - American Time Use Survey (2009-2019).csv
-   Link: <https://github.com/owid/owid-datasets/blob/master/datasets/Who%20Americans%20spend%20time%20with%20-%20American%20Time%20Use%20Survey%20(2009-2019)/Who%20Americans%20spend%20time%20with%20-%20American%20Time%20Use%20Survey%20(2009-2019).csv>

d)  Come up with a (research) question you want to answer with the data and briefly explain why you believe this is an interesting question within one sentence. It should be a question that can be answered with the dataset and using R.

I am interested in how age is related to how much time Americans spend with their friends.

e)  Use R to answer your chosen question.

```{r}
american_time <- read_csv("data/Who Americans spend time with - American Time Use Survey (2009-2019).csv")

american_time_clean <- american_time %>% 
  rename(friends = `Time spent with friends, by age of respondent (United States)`, 
         year = Year)
```


```{r}
cor(american_time_clean$year, american_time_clean$friends)
```


```{r}
model <- lm(friends ~ year, american_time_clean)

summary(model)
```

According to the data, there is a highly significant negative relationship between Americans' age and the amount of time they spend with friends in a day. On average, this amount of time decreases by about 1 minute each year.

f)  Create a meaningful plot / figure with the dataset. Make sure to provide a figure caption (via the chunk options / Rmarkdown) and correctly label the figure.

```{r, fig.cap = "Minutes in a day Americans spend with their friends across their lives"}
ggplot(data = american_time_clean,
       aes(x = year, y = friends)) + 
  geom_line() + 
  labs(x = "age", y = "minutes of day spent with friends") + 
  theme_minimal()
```


## Final Note

Make sure to push all your commits and changes to GitHub before submittining the exercise sheet.
